### 2014-11-10 stat374/hw2.tex
Done homework 5

@@

        \end{itemize}
    \item[3.]
        Starting with the bias:
        \begin{align*}
            bias(\hat p_n(x_0,y_0)) &= p(x_0,y_0) - \E(\hat p_n(x_0,y_0)) \\
                                    &= p(x_0,y_0) - \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(q,t)\,dt\,dq \\ 
                                    &= \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)(p(q,t)-p(x_0,y_0))\,dt\,dq \\ 
                                    &\leq \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)L(\lvert t - x_0 \rvert^\beta +\lvert q - y_0 \rvert^\beta )\,dt\,dq \\ 
                                    &\leq \iint K\left(t\right)K\left(q\right)L(\lvert ht \rvert^\beta +\lvert hq \rvert^\beta )\,dt\,dq \\ 
                                    &\leq 4Lh^\beta\iint_{\R^+\times\R^+} K\left(t\right)K\left(q\right)(t^\beta + q^\beta )\,dt\,dq \\ 
                                    &\leq h^\beta c_1\\ 
        \end{align*}
        Since each $X_i$ is iid and the kernel estimator is linear, we can decompose the variance:
        \begin{align*}
            \var\left(\hat p_n(x_0,y_0)\right) &= \frac{1}{nh^4}\var\left(K\left(\frac{X_i-x_0}{h}\right)K\left(\frac{Y_i-y_0}{h}\right)\right) \\
                                               &= \frac{1}{nh^4}\iint \left( K\left(\frac{x-x_0}{h}\right)K\left(\frac{y-y_0}{h}\right) - \iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(q,t)\,dt\,dq\right)^2 p(x,y)\,dx\,dy \\
        \end{align*}
    \item[4.]
        \begin{itemize}
            \item[a)]
                Here are the histograms of the variable densities. The number of bins was chosen between 1 and 200 based on cross validation score.\\
                \includegraphics[width=9cm]{hw2_4a_Ba_hist}
                \includegraphics[width=9cm]{hw2_4a_Ca_hist} \\ 
                \includegraphics[width=9cm]{hw2_4a_Fe_hist} 
                \includegraphics[width=9cm]{hw2_4a_K_hist}   \\
                \includegraphics[width=9cm]{hw2_4a_Mg_hist}
                \includegraphics[width=9cm]{hw2_4a_Na_hist}  \\
                \includegraphics[width=9cm]{hw2_4a_RI_hist} 
                \includegraphics[width=9cm]{hw2_4a_Si_hist}  \\
            \item[b)]
                These are the density estimates with bandwidth picked by cross validation. The Gaussian Kernel's bandwidth was picked with R's built in unbiased cross validation tool, while the other two I calculated manually on a series of points around the optimal Gaussian bandwidth, picking the smallest estimated risk. In general, the Gaussian seemed to go with the higher bandwidth, producing the smoother estimate, but that may be in part due to a failure in my implementation of cross validation. Certainly it is smoother than the other two kernels by construction, since they are $0$ except over a closed interval. 
                \includegraphics[width=9cm]{hw2_4b_Ba_den}
                \includegraphics[width=9cm]{hw2_4b_Ca_den} \\ 
                \includegraphics[width=9cm]{hw2_4b_Fe_den} 
                \includegraphics[width=9cm]{hw2_4b_K_den}   \\
                \includegraphics[width=9cm]{hw2_4b_Mg_den}
                \includegraphics[width=9cm]{hw2_4b_Na_den}  \\
                \includegraphics[width=9cm]{hw2_4b_RI_den} 
                \includegraphics[width=9cm]{hw2_4b_Si_den}  \\
            \item[c)]
                Cross validation seems to do poorly with variables with one or a few large point masses. We see this several times with a large number of points being at $0$. This is because cross validation rewards the estimator for having an extremely sharp peak around these point masses, even though the rest of the distribution suffers. One might view these as almost two separate distributions. A better way to estimate these sort of variables would be to predict the point mass separately, remove those points, and then run the density estimate on the rest of the distribution.

>>>>

                I ran a generalized additive model with just smoothed functions for daylabel, humidity, temp, windspeed, and hour. The end result was that I got a RMSLE of $1352.043$ for registered customers and $1326.217$ for casual customers.
                For the solution data set I turned in, I predicted the total counts with an additive linear model predicting the log of the total counts with smoothed daylabel, humidity, temp, wind speed, hour, and categorical of working day, and weather interacted with season.
    \item[3.]
        Starting with the bias:
            bias(\hat p_n(x_0,y_0)) &= p(x_0,y_0) - \E(\hat p_n(x_0,y_0)) \\
                                    &= p(x_0,y_0) - \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(q,t)\,dt\,dq \\ 
                                    &= \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)(p(q,t)-p(x_0,y_0))\,dt\,dq \\ 
                                    &\leq \frac{1}{h^2}\iint K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)L(\lvert t - x_0 \rvert^\beta +\lvert q - y_0 \rvert^\beta )\,dt\,dq \\ 
                                    &\leq \iint K\left(t\right)K\left(q\right)L(\lvert ht \rvert^\beta +\lvert hq \rvert^\beta )\,dt\,dq \\ 
                                    &\leq 4Lh^\beta\iint_{\R^+\times\R^+} K\left(t\right)K\left(q\right)(t^\beta + q^\beta )\,dt\,dq \\ 
                                    &\leq h^\beta c_1\\ 
        Since each $X_i$ is iid and the kernel estimator is linear, we can decompose the variance:
            \var\left(\hat p_n(x_0,y_0)\right) &= \frac{1}{nh^4}\var\left(K\left(\frac{X_i-x_0}{h}\right)K\left(\frac{Y_i-y_0}{h}\right)\right) \\
                                               &= \frac{1}{nh^4} \E\left(\left(\iint K\left(\frac{X-x_0}{h}\right)K\left(\frac{Y-y_0}{h}\right)p(t,q) - K\left(\frac{x_0-t}{h}\right)K\left(\frac{y_0-q}{h}\right)p(t,q)\,dt\,dq\right)\right)^2 \\
                                               &\leq \frac{1}{nh^4} \E\left(\left(\iint K\left(\frac{X-t}{h}\right)K\left(\frac{Y-q}{h}\right)p(t,q)\,dt\,dq\right)\right)^2 \\
        And here is where I am stuck. However, if we had an expression for variance, then the mean squared error would be 
        \[bias(\hat p_n(x_0,y_0))^2 + \var\left(\hat p_n(x_0,y_0)\right) \leq h^{2\beta}c_1 + \var\left(\hat p_n(x_0,y_0)\right)\]
        Differentiating the right side of that with respect to $h$ and setting it to $0$, we would find the optimal $h$. That would have a leading term of the order $h^{2\beta-1}$, at least assuming the bias has the larger $h$ power. Then, we would expect to see $h$ shrink with $cn^\frac{-1}{2\beta+1}$., dictating $R$ converge at a rate of $O(n^\frac{2\beta}{2\beta+1})$. 
    \item[4.]
                Here are the histograms of the variable densities. The number of bins was chosen between 1 and 200 based on cross validation score.\\
                \includegraphics[width=9cm]{hw2_4a_Ba_hist}
                \includegraphics[width=9cm]{hw2_4a_Ca_hist} \\ 
                \includegraphics[width=9cm]{hw2_4a_Fe_hist} 
                \includegraphics[width=9cm]{hw2_4a_K_hist}   \\
                \includegraphics[width=9cm]{hw2_4a_Mg_hist}
                \includegraphics[width=9cm]{hw2_4a_Na_hist}  \\
                \includegraphics[width=9cm]{hw2_4a_RI_hist} 
                \includegraphics[width=9cm]{hw2_4a_Si_hist}  \\

@@

                These are the density estimates with bandwidth picked by cross validation. The Gaussian Kernel's bandwidth was picked with R's built in unbiased cross validation tool, while the other two I calculated manually on a series of points around the optimal Gaussian bandwidth, picking the smallest estimated risk. In general, the Gaussian seemed to go with the higher bandwidth, producing the smoother estimate, but that may be in part due to a failure in my implementation of cross validation. Certainly it is smoother than the other two kernels by construction, since they are $0$ except over a closed interval. 
                \includegraphics[width=9cm]{hw2_4b_Ba_den}
                \includegraphics[width=9cm]{hw2_4b_Ca_den} \\ 
                \includegraphics[width=9cm]{hw2_4b_Fe_den} 
                \includegraphics[width=9cm]{hw2_4b_K_den}   \\
                \includegraphics[width=9cm]{hw2_4b_Mg_den}
                \includegraphics[width=9cm]{hw2_4b_Na_den}  \\
                \includegraphics[width=9cm]{hw2_4b_RI_den} 
                \includegraphics[width=9cm]{hw2_4b_Si_den}  \\
                Cross validation seems to do poorly with variables with one or a few large point masses. We see this several times with a large number of points being at $0$. This is because cross validation rewards the estimator for having an extremely sharp peak around these point masses, even though the rest of the distribution suffers. One might view these as almost two separate distributions. A better way to estimate these sort of variables would be to predict the point mass separately, remove those points, and then run the density estimate on the rest of the distribution.
